Resource Buying Games

### 2017-02-28 alg.tex
proceeded on paper, added algorithm on separated pdf for transparency print

\documentclass{beamer}
\usepackage{pgfplots}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage{cancel}
\usepackage{stmaryrd}
\usepackage{url}
\usetheme{metropolis}
\title{Resource Buying Games}
\author[C. Welzel]{Christoph Welzel}

\newcommand{\tupel}[1]{\left(#1\right)}
\newcommand{\set}[1]{\left\{#1\right\}}

\DeclareMathOperator{\ex}{ex}

\begin{document}
\input{tikz/settings}
\maketitle

\begin{frame}
  \frametitle{Introduction}
    \begin{itemize}
      \item Resources: Nodes/edges in a graph, Machines to schedule jobs on,
        \dots
      \item Players: jointly buy resources to meet individual
        (potentially overlapping) demands
      \item<2-> Every player contributes to buy certain resources
      \item<3-> Payoff for a player depends on what she payed to meet her demands
    \end{itemize}
    \uncover<4->{\begin{definition}[Pure Nash Equilibrium(informal)]
      A \emph{Pure Nash Equilibrium (PNE)} is given if for every player holds
      that, given all other players stick to their strategies, she cannot alter
      her behaviour to achieve a better payoff for herself.
    \end{definition}}
\end{frame}

\section{Basic Definitions}
\begin{frame}
  \frametitle{Definition: Congestion Model}
  \begin{equation*}
    \mathcal{M} = \tupel{N, E, \mathcal{S}, \left(d_{i}\right)_{i\in N},
    \left(c_{e}\right)_{e\in E}}
  \end{equation*}
  \vspace{-1cm}
  \begin{itemize}
    \item $N$: set of players
    \item<2-> $E$: set of resources
    \item<3-> $\mathcal{S} = \times_{i\in N}\mathcal{S}_{i}$ with $\mathcal{S}_{i}$
      set of desired resources $S_{i}\subseteq E$: set of configurations
    \item<4-> $d_{i}$: demand of player $i$ for her resource
      \begin{itemize}
        \item<4-> $d_{i} = 1$ for an unweighted game
        \item<4-> $\ell_{e}(S) = \sum_{i\in N:e\in S_{i}}d_{i}$: load of $e$
          in configuration $S\in\mathcal{S}$
      \end{itemize}
    \item<5-> $c_{e}:\mathbb{N}\rightarrow\mathbb{N}$ with
      $c_{e}(S) = c_{e}(\ell_{e}(S))$: cost for the resource $e$ under the load
      of $S$
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Definition: Weighted Resource Buying Game}
  \begin{equation*}
    \mathcal{M} = \tupel{N, E, \mathcal{S}, \left(d_{i}\right)_{i\in N},
    \left(c_{e}\right)_{e\in E}}
  \end{equation*}
  \begin{equation*}
    \leadsto\mathcal{G} = \tupel{N, \mathcal{S}\times\mathcal{P}, \pi}
  \end{equation*}
  \vspace{-1cm}
  \begin{itemize}
    \item<2-> $\mathcal{P}=\times_{i\in N} P_{i}$ with
      $P_{i} = \mathbb{R}^{|E|}_{+}$: payment of player $i$
    \item<3-> $e\in E$ \emph{is bought under $S\in \mathcal{S}$} if
      $\sum_{i\in N}p_{i}^{e} \geq c_{e}(S)$
    \item<4-> $\pi_{i}(S) =
      \begin{cases}
        \sum_{e\in E}p_{i}^{e} &\textit{if }S_{i}\textit{ is bought}\\
        \infty &\textit{otherwise}
      \end{cases}$: {private cost of player $i$}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Restricted Setup}
  \begin{itemize}
    \item \emph{unweighted} games
    \item<2-> marginally non-increasing cost-function $c_{e}$:
      \begin{equation*}
        c_{e}(x + \delta) - c_{e}(x) \geq c_{e}(y + \delta) - c_{e}(y)
        \;\;x\leq y, \delta \in \mathbb{N}
      \end{equation*}
      \begin{center}
        \begin{tikzpicture}
          \begin{axis}[
              xlabel=$x$,
              ylabel=$c_{e}(x)$,
              smooth,
              width=5cm,
              height=3cm,
              ticks=none,
              no markers
            ]
              \addplot{ln(x)};
          \end{axis}
        \end{tikzpicture}
      \end{center}
    \item<3-> restrict players demands to a specific mathematical structure:
      Matroids
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Definition: Matroid}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{equation*}
      M = (E, B)
    \end{equation*}
    \vspace{-0.8cm}
    \begin{itemize}
      \item $E$: ground-set
      \item $B$: set of bases
      \item \emph{basis exchange property}:
        \vspace{-1cm}
        \begin{center}
          \begin{align*}
            \forall &X, Y\in B:\\
            &\forall x\in (X\setminus Y)\\
            &\exists y\in (Y\setminus X):\\
            &\;(X\setminus \set{x})\cup\{y\}\in B
          \end{align*}
        \end{center}
    \end{itemize}
    \column{0.5\textwidth}
    \uncover<2->{Spanning trees:}
    \begin{center}
      \uncover<2->{\resizebox{0.8\textwidth}{!}{\input{tikz/graphp1}}\\[0.5cm]}
      \uncover<3->{\resizebox{0.4\textwidth}{!}{\input{tikz/graphp1S1}}}
      \uncover<4->{\resizebox{0.4\textwidth}{!}{\input{tikz/graphp1S2}}}
      \uncover<5->{\resizebox{0.4\textwidth}{!}{\input{tikz/graphp1S3}}}
    \end{center}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Cuts in Matroids}
  \begin{columns}
    \column{0.7\textwidth}
    $\mathcal{M} = \tupel{E, \mathcal{B}}$
    \begin{itemize}
      \item \emph{Cut}: inclusion-wise minimal $C\subseteq E$ such that
        $E\setminus C$ does not contain a basis of $\mathcal{M}$
      \item[$\leadsto$] $\mathcal{C}\tupel{\mathcal{M}} =
        \set{C\subseteq E\middle| C\text{ is a cut in } \mathcal{M}}$
    \end{itemize}
    \column{0.3\textwidth}
    \resizebox{0.8\textwidth}{!}{\input{tikz/graphp1}}
  \end{columns}
  \vspace{0.7cm}
  \begin{columns}
    \column{0.33\textwidth}
    \uncover<2->{\resizebox{0.7\textwidth}{!}{\input{tikz/graphp1Cut1}}\\[0.3cm]}
    \uncover<3->{\resizebox{0.7\textwidth}{!}{\input{tikz/graphp1Cut2}}}
    \column{0.33\textwidth}
    \uncover<4->{\resizebox{0.7\textwidth}{!}{\input{tikz/graphp1Cut3}}}\\[0.3cm]
    \uncover<5->{\resizebox{0.7\textwidth}{!}{\input{tikz/graphp1Cut4}}}
    \column{0.33\textwidth}
    \uncover<6->{\resizebox{0.7\textwidth}{!}{\input{tikz/graphp1Cut5}}}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Operations on Matroid $\mathcal{M} = \tupel{E, \mathcal{B}}$}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{center}
      Contraction:
    \end{center}
    \vspace{-0.6cm}
    \begin{align*}
      &\mathcal{M}/e = \tupel{E\setminus\set{e}, \mathcal{B}/e}\text{ with}\\
      &\mathcal{B}/e = \set{B\subseteq (E\setminus\set{e})\middle| B\cup\set{e}\in\mathcal{B}}
    \end{align*}
    \column<4->{0.5\textwidth}
    \begin{center}
      Deletion:
    \end{center}
    \vspace{-0.6cm}
    \begin{align*}
      &\mathcal{M}\setminus e = \tupel{E\setminus\set{e},\mathcal{B}\setminus \set{e}}\text{ with}\\
      &\mathcal{B}\setminus e = \set{B\subseteq E\setminus\set{e}\middle|B\in\mathcal{B}}
    \end{align*}
  \end{columns}
  \begin{center}
    \begin{tikzpicture}
      \uncover<2->{\node (G) {\input{tikz/graphp1}};
      \node[right=1 of G] (leadsto) {\Huge$\leadsto$};}
      \uncover<3>{\node[right=1 of leadsto] (G') {\input{tikz/graphp1Cont}};}
      \uncover<5>{\node[right=1 of leadsto] (G'') {\input{tikz/graphp1Del}};}
    \end{tikzpicture}
  \end{center}
\end{frame}

\section{Algorithm: PNE in Unweighted Matroid Games}
\subsection{Algorithm}
\begin{frame}
  \frametitle{Algorithm}
  \begin{tabular}{|lll|}
    \hline
    (i) & Init: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item \parbox{0.7\textwidth}{every player: empty working basis, complete ground-set, complete set of valid result bases}
          \item every resource: cost for next load, current load
        \end{itemize}
      }\\\hline
    (ii) & Find: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item inclusion-min (necessary cost)-max cut $C$
          \item minimal cost resource $e$ in $C$
          \item owner $i$ of $C$
        \end{itemize}
      }\\\hline
    (iii) & Pick: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item $i$: add $e$ to working basis, contract matroid to $e$
          \item $e$: update marginal cost and increase load
          \item all player: delete elements $C\setminus\set{e}$ from matroid
        \end{itemize}
      }\\\hline
    (iv) & Break: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item exists player with incomplete basis: goto (ii)
          \item otherwise: return chosen bases and payments
        \end{itemize}
      }\\\hline
  \end{tabular}
\end{frame}

\begin{frame}
  \frametitle{Example: Graph}
  \begin{columns}
    \column{0.5\textwidth}
    Player 1:\\
    \vspace{-0.2cm}
    \begin{center}
      \resizebox{0.6\textwidth}{!}{\input{tikz/graphp1labeled}}
    \end{center}
    \column{0.5\textwidth}
    Player 2:\\
    \vspace{-0.5cm}
    \begin{center}
      \resizebox{0.6\textwidth}{!}{\input{tikz/graphp2labeled}}
    \end{center}
  \end{columns}
  \begin{center}
    \resizebox{0.35\textwidth}{!}{\input{tikz/graphboth}}
  \end{center}
\end{frame}

\newlength{\cutscaling}
\setlength{\cutscaling}{0.17\textwidth}
\begin{frame}
  \frametitle{Example: Algorithm}
  \begin{columns}
    % left column:
    \column{0.3\textwidth}
    \resizebox{\textwidth}{!}{\alt<1-6>{\input{tikz/graphp1}}{\alt<7>{\input{tikz/graphp1Cont}}{\input{tikz/graphp1ContDel}}}}
    \begin{columns}
      \column{0.5\textwidth}
      \resizebox{\cutscaling}{!}{\alt<1-2>{\input{tikz/graphp1Cut1}}{\input{tikz/graphp1Cut1value}}}
      \resizebox{\cutscaling}{!}{\alt<1-3>{\input{tikz/graphp1Cut2}}{\input{tikz/graphp1Cut2value}}}
      \resizebox{\cutscaling}{!}{\alt<1-3>{\input{tikz/graphp1Cut3}}{\input{tikz/graphp1Cut3value}}}
      \column{0.5\textwidth}
      \resizebox{\cutscaling}{!}{\alt<1-3>{\input{tikz/graphp1Cut4}}{\input{tikz/graphp1Cut4value}}}
      \resizebox{\cutscaling}{!}{\alt<1-3>{\input{tikz/graphp1Cut5}}{\input{tikz/graphp1Cut5value}}}
    \end{columns}

    % middle column:
    \column{0.4\textwidth}
    \resizebox{\textwidth}{!}{\alt<1-8>{\input{tikz/graphbothCosts}}{\input{tikz/graphbothCostsUpd}}}
    \begin{center}
      \uncover<5->{\alt<5>{\resizebox{0.7\textwidth}{!}{\input{tikz/graphp1Cut3}}}{\resizebox{0.7\textwidth}{!}{\input{tikz/graphp1Cut3marked}}}}
    \end{center}

    % right column:
    \column{0.3\textwidth}
    \resizebox{\textwidth}{!}{\alt<1-7>{\input{tikz/graphp2}}{\input{tikz/graphp2Del}}}
    \begin{columns}
      \column{0.5\textwidth}
      \resizebox{\cutscaling}{!}{\alt<1-3>{\input{tikz/graphp2Cut1}}{\input{tikz/graphp2Cut1value}}}
      \resizebox{\cutscaling}{!}{\alt<1-3>{\input{tikz/graphp2Cut2}}{\input{tikz/graphp2Cut2value}}}
      \resizebox{\cutscaling}{!}{\alt<1>{\input{tikz/graphp2Cut3}}{\input{tikz/graphp2Cut3deleted}}}
      \column{0.5\textwidth}
      \resizebox{\cutscaling}{!}{\alt<1>{\input{tikz/graphp2Cut4}}{\input{tikz/graphp2Cut4deleted}}}
      \resizebox{\cutscaling}{!}{\alt<1>{\input{tikz/graphp2Cut5}}{\input{tikz/graphp2Cut5deleted}}}
      \resizebox{\cutscaling}{!}{\alt<1-3>{\input{tikz/graphp2Cut6}}{\input{tikz/graphp2Cut6value}}}
    \end{columns}
  \end{columns}
  \uncover<9>{ }
\end{frame}

\begin{frame}
  \frametitle{Soundness}
  \begin{itemize}
    \item Termination:
      \begin{itemize}
        \item<2-> inductively applying contraction operations lead to basis
        \item<3-> deletion cannot \enquote{loose} bases for other players since
          chosen cut is inclusion-wise minimal
      \end{itemize}
    \item Correctness
      \begin{itemize}
        \item<4-> Lemma: cost of the chosen element decreases monotonically in each
          iteration
        \item<5-> assume there is a beneficial alternative basis
          $\hat{B} = (B_{i}\setminus\set{g})\cup{f}$
          \begin{itemize}
            \item<6-> consider iteration $k$ in which $g$ is added to basis of $i$:
            \item<7-> if $f$ is in the corresonponding cut its cost never changes
              since it is deleted from every matroid $\rightarrow$ $f$ is not
              in the cut
            \item<8-> but $f$ and $g$ \enquote{open} same dimensionality, thus $f$
              is not left in contracted matroid of $i$
            \item<9-> $f$ was deleted before (in this iteration the cost of the
              chosen element was lower than for $f$ in contradiction to lemma)
          \end{itemize}
      \end{itemize}
  \end{itemize}
\end{frame}

\section{Weighted Matroid Games}
\begin{frame}
  \frametitle{Recall: Congestion Model}
  \begin{equation*}
    \mathcal{M} = \tupel{N, E, \mathcal{B}, \left(d_{i}\right)_{i\in N},
    \left(c_{e}\right)_{e\in E}}
  \end{equation*}
  \vspace{-1cm}
  \begin{itemize}
    \item before: $d_{i} = 1$ for all $i$
    \item<2-> now: $d_{i}\in\mathbb{N}_{>0}$ 
    \item<3-> auxiliary definitions:
      \begin{itemize}
        \item<4-> $N_{e}\tupel{B} = \set{i\in N\middle|e\in B_{i}}$
        \item<5-> for a fixed $B$: $\ex_{i}\tupel{e} =
          \set{f\in E\setminus\set{e}\middle|
          (B_{i}\setminus\set{e})\cup\set{f}\in\mathcal{B}_{i}}$: set of resources
          that can replace $e$ for player $i$
        \item<6-> $F = \bigcup_{i\in N}\bigcap_{B\in \mathcal{B}_{i}} B$:
          set of obligatory elements for at least one player
        \item<7-> $\Delta_{i}\tupel{B;e\rightarrow f} =
          c_{f}\tupel{\ell_{f}\tupel{B_{i}\cup\set{f}\setminus\set{e}}} -
          c_{f}\tupel{\ell_{f}\tupel{B}}$: marginal cost to buy resource $f$ if
          $i$ switches
        \item<8-> $\Delta_{i}^{e} = \min_{f\in\ex_{i}(e)}\set{\Delta_{i}(B;e\rightarrow f)}$:
          minimal marginal cost for exchanging $e$
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Characterisation of Pure Nash Equilibria}
  \begin{theorem}
    For a weighted matroid resource buying game
    $\tupel{N, E, \mathcal{B}, d, c}$ and $B\in\mathcal{B}$ the following
    holds: there is a payment vector $p$ such that $\tupel{B,p}$ is a
    PNE if and only if
    \begin{equation*}
      c_{e}(B) \leq \sum_{i\in N_{e}(B)}\Delta_{i}^{e}(B)
      \textit{ for all }e\in E\setminus F
    \end{equation*}
  \end{theorem}
  \begin{proof}
    if: \uncover<2->{$p_{i}^{e} =
      \begin{cases}
        \frac{\Delta_{i}^{e}(B)}{\sum_{j\in N_{e}(B)}\Delta_{j}^{e}(B)}\cdot
        c_{e}(B)&\text{ if } \sum_{j\in N_{e}(B)}\Delta_{j}^{e}(B) > 0\\
        0 &\text{ otherwise}
      \end{cases}$\\
      for all $e\in E\setminus F$ and $i\in N_{e}(B)$\\}
    only if: \uncover<3->{for all $i$ and $e$: $p_{i}^{e} \leq \Delta_{i}^{e}(B)$
      and $c_{e}(B) = \sum_{i\in N_{e}(B)}p_{i}^{e}$}
  \end{proof}
\end{frame}

\begin{frame}
  \frametitle{Existence of Pure Nash Equilibria}
  \begin{theorem}
    Every weighted matroid resource buying game with marginally non-increasing
    cost functions possesses a PNE.
  \end{theorem}
  \begin{itemize}
    \item<2-> socially optimal configuration profile: $B\in\mathcal{B}$ such that
      $\sum_{e\in E}c_{e}(\ell_{e}(B))$ is minimal
    \item<3-> for socially optimal configuration profile $B$ holds:
      \begin{equation*}
        c_{e}(B)\leq \sum_{i\in N_{e}(B)}\Delta_{i}^{e}(B)
        \text{ for all }e\in E\setminus F
      \end{equation*}
    \item<4-> for $B$ exists $p$ s.t. $\tupel{B,p}$ is a PNE
    \item<5-> (but finding socially optimal configuration profiles is NP-hard)
  \end{itemize}
\end{frame}

\section{Non-Matroid Strategy Spaces}
\begin{frame}
\frametitle{Consequences of using Non-Matroid Strategy Spaces}
  \begin{lemma}
    If $\mathcal{S}\subseteq 2^{E}$ is a non-matroid anti-chain, then there
    exist $X,Y\in\mathcal{S}$ and $\set{a,b,c}\subseteq X\cup Y$ such that each
    set $Z\in\mathcal{S}$ with $Z\subseteq(X\cup Y)\setminus\set{a}$ contains
    $b$ and $c$. (Especially, there is $S\in\mathcal{S}$ with
    $S\subseteq X\cup Y$ and $a\in S$ but $b,c\notin S$.)
  \end{lemma}
  \uncover<2->{\begin{theorem}
    For every non-matroid anti-chain $\mathcal{S}$ on a set of resource $E$,
    there exists a weighted two-player resource buying game
    $G = \tupel{\tilde{E}, \tupel{\mathcal{S}_{1} \times
    \mathcal{S}_{2}}\times \mathcal{P}, \pi}$ having marginally non-increasing
    cost functions, whose strategy spaces $\mathcal{S}_{1}$ and
    $\mathcal{S}_{2}$ are both isomophic to $\mathcal{S}$, so that $G$ does not
    possess a pure Nash equilibrium.
  \end{theorem}}
\end{frame}

\begin{frame}
  \frametitle{Construction of games without PNE}
   \begin{columns}
     \column{0.3\textwidth}
       \begin{itemize}
         \item $c_{x}(t) = t$
         \item $c_{y}(t) = 5\frac{1}{2}$
         \item $c_{z}(t) = 4$
         \item $d_{1} = 5$
         \item $d_{2} = 4$
       \end{itemize}
     \column{0.7\textwidth}
       \begin{center}
         \resizebox{\textwidth}{!}{\input{tikz/gamediagram}}
       \end{center}
   \end{columns}
   \begin{itemize}
     \item<2-> assume PNE $\tupel{Z^{\ast} = \tupel{Z_{1}^{\ast}, Z_{2}^{\ast}},
       p^{\ast} = \tupel{p_{1}^{\ast}, p_{2}^{\ast}}}$
     \item<3-> $Z_{i}^{\ast}\subseteq\tupel{X_{i}\cup Y_{i}}$
     \item<4-> $x\not\in Z_{1}^{\ast}\Rightarrow \set{y,z}\subseteq Z_{1}^{\ast}$
     \item<5-> $y\not\in Z_{2}^{\ast}\Rightarrow \set{x,z}\subseteq Z_{2}^{\ast}$
     \item<6->[$\Rightarrow$] neiter $x\in Z_{1}^{\ast}$ nor $x\not\in Z_{1}^{\ast}$ $\lightning$
   \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Non-Decreasing Marginal Cost Functions}
  \begin{columns}
    \column{0.5\textwidth}
      Non-Increasing Marginal Cost Functions:
      \begin{center}
        \begin{tikzpicture}
          \begin{axis}[
              xlabel=$x$,
              ylabel=$c_{e}(x)$,
              smooth,
              width=5cm,
              height=3cm,
              ticks=none,
              no markers
            ]
              \addplot{ln(x)};
          \end{axis}
        \end{tikzpicture}
      \end{center}
    \column{0.5\textwidth}
      Non-Decreasing Marginal Cost Functions:
      \begin{center}
        \begin{tikzpicture}
          \begin{axis}[
              xlabel=$x$,
              xmin=1,
              xmax=4,
              ylabel=$c_{e}(x)$,
              smooth,
              width=5cm,
              height=3cm,
              ticks=none,
              no markers
            ]
              \addplot{5*exp(x)};
          \end{axis}
        \end{tikzpicture}
      \end{center}
  \end{columns}
  \uncover<2->{\begin{theorem}
    Let $G$ be a weighted resource buying game with non-decreasing marginal
    costs and let $S^{\ast}$ be a socially optimal solution. We can compute
    (so called) marginal cost pricing vectors which induce a PNE.
  \end{theorem}}
\end{frame}

\section{Conclusion}
\begin{frame}
  \frametitle{Results}
  \begin{itemize}
    \item Resource Buying Games
    \item<2-> Matroids:
      \parbox{6cm}{
        \begin{itemize}
          \item Operations: Contraction, Deletion
          \item Cuts
        \end{itemize}
      }
    \item<3-> Games:
      \parbox{8cm}{
        \begin{itemize}
          \item \parbox{6cm}{
              \begin{tabular}{|l|ll|}
                \hline
                Non-increasing marginal & Matroid & No Matroid\\\hline
                unweighted & Algorithm & \\
                weighted & PNE & No PNE\\\hline
                Non-decreasing marginal & PNE & \\\hline
              \end{tabular}
            }
        \end{itemize}
      }
    \item<4-> Slides: \url{github.com/weltoph/sem-alg-gametheory}
  \end{itemize}
  \uncover<5->{\centering Thank you for your attention!}
\end{frame}

\end{document}

>>>>

\documentclass{scrartcl}
\newcommand{\tupel}[1]{\left(#1\right)}
\newcommand{\set}[1]{\left\{#1\right\}}
\begin{document}
  \begin{tabular}{|lll|}
    \hline
    (i) & Init: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item \parbox{0.7\textwidth}{every player: empty working basis,
            complete ground-set, complete set of valid result bases}
          \item every resource: cost for next load, current load
        \end{itemize}
      }\\\hline
    (ii) & Find: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item inclusion-min (necessary cost)-max cut $C$
          \item minimal cost resource $e$ in $C$
          \item owner $i$ of $C$
        \end{itemize}
      }\\\hline
    (iii) & Pick: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item $i$: add $e$ to working basis, contract matroid to $e$
          \item $e$: update marginal cost and increase load
          \item all player: delete elements $C\setminus\set{e}$ from matroid
        \end{itemize}
      }\\\hline
    (iv) & Break: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item exists player with incomplete basis: goto (ii)
          \item otherwise: return chosen bases and payments
        \end{itemize}
      }\\\hline
  \end{tabular}
\end{document}

### 2017-02-28 paper.tex
proceeded on paper, added algorithm on separated pdf for transparency print

We use matroids as a generalisation of the concept of linear independence.
Matroids can be defined as a ground-set and the family of linear
independent sets of this ground-set. This family has to satisfy three
properties, i.e.
\begin{enumerate}
  \item the empty set is linear independent,
  \item all subset of a linear independent set are linear independent and
  \item for two linear independent sets we can find an element of the larger
    set that forms a linear independent set when added to the smaller set.
\end{enumerate}
We use an equivalent definition of matroids which in a way captures the idea of
having a vector space by giving instead of linear independent sets the bases of
the ground-set.
\begin{defi}
  The structure $\tupel{E, \mathcal{B}}$ with $\mathcal{B}\subseteq 2^{E}$ is
  a matroid if
  \begin{enumerate}
    \item $\mathcal{B}\neq\emptyset$
    \item let $B,B'\in\mathcal{B}$ and for every $x\in B$ there is $y\in B'$
      s. t. $(B\setminus\set{x})\cup\set{y}\in\mathcal{B}$
  \end{enumerate}
\end{defi}
In analogy to vector spaces we consider every element of a basis to open access
to a dimension. We illustrate this concept with the following example: given an
undirected connected graph $G = \tupel{V, E}$ we will argue that
$\tupel{E, \mathcal{S}}$ with $\mathcal{S}$ as edge-sets of spanning trees
in $G$ forms a matroid. In order to meet the requirement that $\mathcal{S}$
must not be empty we restrict $G$ to those graphs with at least two distinct
vertices. Furthermore, given two spanning trees $\tupel{V, S_{1}}$ and
$\tupel{V, S_{2}}$ of $G$ then removing an edge from $S_{1}$ yields a subgraph
of $G$ that is separated into two connected components. Since $S_{2}$ forms a
spanning tree as well there must be an edge connecting these components which
can be used to again obtain a spanning tree\footnote{see Figure
\ref{fig:spanning} for an illustration}. Consider the algorithm of Prim
which computes spanning trees by iteratively adding edges which do not close a
cycle: every added edge works towards a basis in $\mathcal{S}$ and opens a
dimensionality by connecting one of the graph nodes.
\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \node (t) {\input{tikz/graphp1}};
      \node[below left=of t] {\input{tikz/graphp1S1}};
      \node[below =of t] {\input{tikz/graphp1S2}};
      \node[below right=of t] {\input{tikz/graphp1S1}};
    \end{tikzpicture}
  \end{center}
  \caption{Graph with its spanning trees}
  \label{fig:spanning}
\end{figure}
\todo{working iteratively towards a basis by prim's algorithm as example to
introduce cuts and contractions and deletions}

>>>>

We use matroids as a generalisation of the concept of linear independence.
Matroids can be defined as a ground-set and the family of linear
independent sets of this ground-set. This family has to satisfy three
properties, i.e.
\begin{enumerate}
  \item the empty set is linear independent,
  \item all subset of a linear independent set are linear independent and
  \item for two linear independent sets we can find an element of the larger
    set that forms a linear independent set when added to the smaller set.
\end{enumerate}
We use an equivalent definition of matroids which in a way captures the idea of
having a vector space by giving instead of linear independent sets the bases of
the ground-set.
\begin{defi}
  The structure $\tupel{E, \mathcal{B}}$ with $\mathcal{B}\subseteq 2^{E}$ is
  a matroid if
  \begin{enumerate}
    \item $\mathcal{B}\neq\emptyset$
    \item let $B,B'\in\mathcal{B}$ and for every $x\in B$ there is $y\in B'$
      s. t. $(B\setminus\set{x})\cup\set{y}\in\mathcal{B}$
  \end{enumerate}
\end{defi}
In analogy to vector spaces we consider every element of a basis to open access
to a dimension. We illustrate this concept with the following example: given an
undirected connected graph $G = \tupel{V, E}$ we will argue that
$\tupel{E, \mathcal{S}}$ with $\mathcal{S}$ as edge-sets of spanning trees
in $G$ forms a matroid. In order to meet the requirement that $\mathcal{S}$
must not be empty we restrict $G$ to those graphs with at least two distinct
vertices. Furthermore, given two spanning trees $\tupel{V, S_{1}}$ and
$\tupel{V, S_{2}}$ of $G$ then removing an edge from $S_{1}$ yields a subgraph
of $G$ that is separated into two connected components. Since $S_{2}$ forms a
spanning tree as well there must be an edge connecting these components which
can be used to again obtain a spanning tree\footnote{see Figure
\ref{fig:spanning} for an illustration}.
\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \node (t) {\input{tikz/graphp1}};
      \node[below left=of t] {\input{tikz/graphp1S1}};
      \node[below =of t] {\input{tikz/graphp1S2}};
      \node[below right=of t] {\input{tikz/graphp1S1}};
    \end{tikzpicture}
  \end{center}
  \caption{Graph with its spanning trees}
  \label{fig:spanning}
\end{figure}
\todo{working iteratively towards a basis by prim's algorithm as example to
introduce cuts and contractions and deletions}
Consider the algorithm of Prim which computes spanning trees by iteratively
adding edges which do not close a cycle: every added edge works towards a basis
in $\mathcal{S}$ and opens a dimensionality by connecting one of the graph
nodes. When an edge is added the result of the algorithm can only compute
certain spanning trees, namely those which contain all edges added up to this
point. Fixing certain elements in matroids and therefore restricting the set of
\enquote{useable} bases is captured by the \emph{contraction} operation on
matroids.
\begin{defi}
  For a matroid $\mathcal{M} = \tupel{E, \mathcal{B}}$ and one $e\in E$ we call
  the matroid $\tupel{E\setminus\set{e}, \set{B\subseteq
  (E\setminus\set{e})\middle| B\cup\set{e}\in\mathcal{B}}}$ the
  \emph{contraction} of $e$ in $\mathcal{M}$, denoted by $\mathcal{M}/e$.
\end{defi}
In a similar manner we introduce the \emph{deletion} operation on matroids to
disregard all bases which use a certain element.
\begin{defi}
  For a matroid $\mathcal{M} = \tupel{E, \mathcal{B}}$ and one $e\in E$ we call
  the matroid $\tupel{E\setminus\set{e}, \set{C\subseteq
  (E\setminus\set{e})\middle| B\in\mathcal{B}}}$ the \emph{deletion} of $e$ in
  $\mathcal{M}$, denoted by $\mathcal{M}\setminus e$.
\end{defi}
Note that this deletion operation is not necessarily well defined for all
elements of the ground-set, e.g. deletion of an edge in the spanning tree
matroid which is a bridge yields an empty set of bases and thus no valid
matroid.
